---
title: "Little_Black_Book_Of_R"
author: "Elisa Carrasco Lanusse"
date: "`r Sys.Date()`"
output: html_document
---

# BASICS -----------------------------------------
```{r}
# Mathematical symbols work as you would expect them to
object_name <- value # Creates an object: <- (option + -)
primes <- c(2, 3, 5, 7, 11, 13) # Combines multiple elements into vector (c())

i_use_snake_case
otherPeopleUseCamelCase
some.people.use.periods
And_aFew.People_RENOUNCEconvention

# Pipes
df |> # Pipes are extra helpful when writing big chunks
  function() 

function_name(argument1 = value1, argument2 = value2, ...) # Function generally have this structure

reprex::reprex()


```

# KEY FUNCTIONS (RUN THIS BITCH) ---------------------------------------
```{r}

# Source Functions
source("/Users/elisacarrasco/Documents/GitHub/learning-r/Functions.R")

# key_packages_fn - These are packages I tend to work with often. We are installing and setting up collectivly.
key_packages_fn <- function() {
  # List of packages
  key_packages <- c("tidyverse", "readr", "dplyr", "ggplot2", "rmarkdown", "xml2", "plotly", "shiny", "ggthemes", "janitor", "styler", "arrow", "ggsci", "rstudioapi", "ggridges", "patchwork", "lvplot", "hexbin")
  # Check if each package is installed, and install it if not
  for (pkg in key_packages) {
    if (!requireNamespace(pkg, quietly = TRUE)) {
      install.packages(pkg)
    }
    # Load the package
    library(pkg, character.only = TRUE)
  }
}

# Call the function
key_packages_fn()




janitor::clean_names() # Great for cleaning up names

```

# SETUP
```{r}
data_frame <- read_csv("name.csv", na = c("N/A", "") | data_frame <- read_csv("https", na = c("N/A", "") # Clears na
                                                                              
# Types of reads
    read_csv2() # Reads semicolon-separated files. These use ; instead of , to separate fields and are common in countries that use , as the decimal marker.

    read_tsv() # Reads tab-delimited files.

    read_delim() # Reads in files with any delimiter, attempting to automatically guess the delimiter if you don’t specify it.

    read_fwf() # Reads fixed-width files. You can specify fields by their widths with fwf_widths() or by their positions with fwf_positions().

    read_table() # Reads a common variation of fixed-width files where columns are separated by white space.

    read_log() #Reads Apache-style log files.

write_csv(df, "df.csv")

# Set working diretory
setwd("/Users/elisacarrasco/Desktop")

# Set directory for data frame and file paths
directory_name <- "/Users/elisacarrasco/Desktop/Name"
name_paths <- list.files(directory_name, pattern = ".filetype", full.names = TRUE)
name_list <- lapply(name_paths, read_filetype)

# Combine all data frames into one and view
name_df <- do.call(rbind, name_list)
print(name_df)
View(name_df)
```

# DATA ENTRY
```{r}
tribble( # A tribble can be entered by hand, defining the columns (variables) and the rows (values)
  ~x, ~y, ~z,
  1, "h", 0.08,
  2, "m", 0.83,
  5, "g", 0.60)


```

# READR --------------------------------------------
```{r}
# Column Types
col_logical() # Reads logicals and real numbrs
col_integer() # Reads integers
col_character() # Reads strings
col_factor() # Create factors
col_date() # Create dates
col_datetime() # Create date-times
col_number() # Ignores non-numeric components, useful for currencies
col_skip() # Skips column so it's not included
cols_only() # Reads only specified columns

write_csv(name, "name.csv") # For writing a file

read_rds() / write_rds() # Store the exact R object that you stored

read_parquet() / write_parquet() # File format that can be shared across languages

tibble(         # Manually set up a tibble
  x = c(1,2,3),
  y = c("h","m", "g"),
  z = c(0.08, 0.83, 0.60)
)

tribble(        # Transported tibble
  ~x, ~y, ~z,
  1, "h", 0.08,
  2,"m", 0.83,
  5, "g", 0.60
)
```

# TIDYVERSE --------------------------------------------
```{r}
install.packages("tidyverse")
library(tidyverse)

# Rules of tidyverse: Each variable is a column; each column is a variable.   Each observation is a row; each row is an observation.  Each value is a cell; each cell is a single value.

pivot_longer(cols = parameter), 
    names_to = "columnname", 
    values_to = "columnname", 
    values_drop_na = TRUE) # cols specifies which columns need to be pivoted, i.e. which columns aren’t variables.   names_to names the variable stored in the column names.   values_to names the variable stored in the cell values.

pivot_wider(
    id_cols = # Which columns are being used,
    names_from = # Where the names are taken from,
    values_from = # Where the data is from
  )
```
- df = data frame
Rules of tidy data:
    Each variable is a column; each column is a variable.
    Each observation is a row; each row is an observation.
    Each value is a cell; each cell is a single value.


# DPLYR ----------------------------------------------
```{r}
ungroup()
flights |>
  filter(dest == "IAH") |> 
  group_by(year, month, day) |> 
  summarize(
    arr_delay = mean(arr_delay, na.rm = TRUE)
  )

|> or %>% (command + shift + m) # This is a pipe, takes the thing on its left and passes it along to the function on its right so that x |> f(y) is equivalent to f(x, y), and x |> f(y) |> g(z) is equivalent to g(f(x, y), z)

# Rows
filter() # Changes which rows are present without changing their order
  # Use greater than >, greater than or equal >=, less than <, less than or equal <=, equal to ==, not equal to !=, and &, or |.
  # %in% Keeps rows where the variable equals one of the values on the right.
arrange() # Changes the order of the rows without changing which are present
  # arrange(desc()) Arranges the data in the column in descending order
distinct() # Finds all the unique rows in a dataset
count() # Counts the number of ocurrences of each row

# Columns
mutates() # Add new columns that are calculated from the existing columns
  # .before and .after can be used to dictate where the new column is placed in relation to an existing column
  # .keep Dictates which columns are displayed. .keep = used specifies that we only keep the columns that were involved
select() # Zoom in on a useful subset using operations based on the names of the variables
  # starts_with("abc"): matches names that begin with “abc”.  ends_with("xyz"): matches names that end with “xyz”.   contains("ijk"): matches names that contain “ijk”.  num_range("x", 1:3): matches x1, x2 and x3.
rename() 
relocate() # 

# Groups
group_by() # Divide dataset into groups meaningful
summarize() 
slice_ # df |> slice_head(n = 1) takes the first row from each group.   df |> slice_tail(n = 1) takes the last row in each group.   df |> slice_min(x, n = 1) takes the row with the smallest value of column x.  df |> slice_max(x, n = 1) takes the row with the largest value of column x.   df |> slice_sample(n = 1) takes one random row.
ungroup()

```
Dplyr structure: The first argument is always a data frame.  The subsequent arguments typically describe which columns to operate on using the variable names (without quotes).   The output is always a new data frame.

# GGPLOT -------------------------------------------
```{r}
## General structure
name_plot <- ggplot(
  dfname, aes(x = xname, y = yname)) +
  geom_graphtype(aes(color = variable_a, shape = variable_b)) + # Sometimes a geom_type will have different usage methods -> geom_smooth(method = lm)
  labs(
    title = "Title",
    subtitle = "Subtitle",
    x = "Variable A (metric)", y = "Variable B (metric)",
    color = "Variable C", shape = "Variable C") +
  scale_color_colorblind() # Accessible colors, might not be necessary if using different shapes

# Types of plots
Bar A (categorical data) <- ggplot(df, aes(x = fct_infreq(variable))) + ## fct_ sets order of categorical variables, infreq is frequency based
  geom_bar() 

Bar B (frequencies) <- ggplot(df, aes(x = variable, fill = variable)) +
  geom_bar(position = "fill")
  # position = "fill" works like stacking, but makes each set of stacked bars the same height. This makes it easier to compare proportions across groups.
  # position = "dodge" places overlapping objects directly beside one another. This makes it easier to compare individual values.

Bar C (relative freq) <- ggplot(df, aes(x = variable, fill = variable)) +
  geom_bar(position = "fill")

Boxplot (categorical and numerical) <- ggplot(df, aes(x = variable, y = variable)) +
  geom_boxplot()

Boxplot (letter values) <- ggplot(df, aes(x = variable, y = variable)) +
  geom_lv(fill = ..LV..)

Density A (numerical data) <- ggplot(df, aes(x = variable)) +
  geom_density(linewidth = number)

Density B (numerical) <- ggplot(df, aes(x = variable, color = variable, fill = variable)) +
  geom_density(alpha = number)
# Alpha changes the opacity

Frequencyplot <- ggplot(df, aes(x = variable_1, y = after_stat(density))) +
  geom_freqpoly(aes(color = variable_2))

Histogram (numerical) <- ggplot(df, aes(x = variable)) +
  geom_histogram(binwidth = size)

Scatterplot (numerical, 3+ variables) <- ggplot(df, aes(x = variable_1, y = variable_2)) +
  geom_point(aes(color = variable_3, shape = variable_4)) +
  geom_jitter() 
# position = "jitter" / geom_jitter() helps with overplotting, especially important for large scales. Can replace geom_point with geom_jitter.
# Alternatives to geom_jitter() are:
  # geom_beeswarm
  # geom_quasirandom

Violinplot <- ggplot(df, aes(x = variable_1, y = variable_2, fill = variable_2, color = variable_2)) +
  geom_violin()

# Saving Plots
ggsave(filename = "filename.png")

# Things can be specified within the aesthetics:
  color = "color_name"
  size = number
  shape = "shape_name"
  stroke = number
  
# Facets
ggplot(df, aes(x = variable_1, y = variable_2)) + 
  geom_point() + 
  facet_wrap(~variable_3) # Splits the plot into subplots based on categorical variable 
# Using facet_wrap cleans up graphs like these into subsets :)

ggplot(df, aes(x = variable_1, y = variable_2)) + 
  geom_point() + 
  facet_grid(variable_3 ~ variable_4 scales = "free") # Facets plot with the combination of two variables, creating a grid. Freeing the scales allows different scales per variables

# Bar charts, histograms, and frequency polygons bin your data and then plot bin counts, the number of points that fall in each bin.

# Smoothers fit a model to your data and then plot predictions from the model.

# Boxplots compute the five-number summary of the distribution and then display that summary as a specially formatted box.

## Coordinate Systems
coord_quickmap() # Sets the aspect ratio correctly for geographic maps

coord_polar() # Polar coordinates reveal interesting connections between a bar chart and a coxcomb chart. REALLY COOL, makes a pie chart but fancy.

## Layered Grammar
ggplot(df = <DATA>) + 
  <GEOM_FUNCTION>(
     mapping = aes(<MAPPINGS>),
     stat = <STAT>, 
     position = <POSITION>
  ) +
  <COORDINATE_FUNCTION> +
  <FACET_FUNCTION>
```

# EXPLORATORY DATA ANALYSIS
```{r}
# Variation (Behavior within variables)
"1. Which values are the most common? Why?",
"2. Which values are rare? Why? Does that match your expectations?",
"3. Can you see any unusual patterns? What might explain them?"

Unusual_Values <- ggplot(df, aes(x = y)) + 
  geom_histogram(binwidth = 0.5) +
  coord_cartesian(ylim = c(0, 50)) # coord_cartesian() zooms in to small values on the axis, (ylim / xlim). Good for finding outliers.

data_set_na_outliers <- df |> 
  mutate(y = if_else(y < number | y > number, NA, y)) # Replacing outliers with NA helps avoid confusion when we want to get rid of them

# Covariation (Behavior between variables)

Covariation_cat_num <- ggplot(df, aes(x = catvariable_1, y = after_stat(density))) + # after_stat lets you calculate certain statistics in order to become one of the visualized variables
  geom_freqpoly(aes(color = variable_2), binwidth = 500, linewidth = 0.75)

Covariation_cat_cat_SIZE <- ggplot(df, aes(x = variable_1, y = variable_2)) +
  geom_count() # Plots the count of each covariation, size guides

Covariation_cat_cat_COLOR <- df |> 
  group_by(variable_1, variable_2) |> 
  count(variable_1, variable_2) |>  
  ggplot(aes(x = variable_1, y = variable_2)) +
  geom_tile(aes(fill = n)) # Plots the count of each covariation, color guides

Covariation_num_num <- ggplot(df, aes(x = variable_1, y = variable_2)) +
  geom_point() # This is especially useful for smaller data sets

Covariation_num_num <- ggplot(df, aes(x = variable_1, y = variable_2)) +
  geom_bin2d() # Useful for large data sets, creates rectangular bins.

Covariation_num_num <- ggplot(df, aes(x = variable_1, y = variable_2)) +
  geom_hex() # Useful for large data sets, creates hexagonal bins.




- after_stat(density) # Density is a standardized count that is calculated form the values. Useful to get clear visualizations in cases like this where the raw count might not be representative.


```
  - EDA involves two types of questions:
    1. What type of variation occurs between variables?
    2. What type of covariation occurs between my variables?



